\LARGE \bf Generalizing Informed Sampling for Asymptotically Optimal Sampling-based Kinodynamic Planning via Markov Chain Monte Carlo

### 2017-09-16 paper.tex
update new plot

Given a trajectory $\gamma_{\text{best}}$ with cost $c_{\text{best}} = c(\gamma_{\text{best}})$ the \emph{informed set}~\Cinf is defined to be all states $x$  which may be on trajectories with lower cost than $c_{\text{best}}$.
Specifically,
$
\Cinf = \{ x \in \calX \mid  
		c ( \gamma^*(x) ) < \cbest \} $~\cite{GSB14}.
Here~$ \gamma^*(x) $ denotes the optimal trajectory  from $ x_s $ to $ x_g $ constrained to pass through $ x $.
Notice that we do not require that~$ \gamma^*(x) $ is collision free.

@@

%Let $T(x_1, x_2)$ denote the minimal time to move the robot from state $x_1$ to $x_2$.
%Furthermore, set
%%
%\begin{equation}
%T(x_s, x, x_g) = T(x_s, x) + T(x, x_g).
%\end{equation}
%%
%Namely, $T(x_s, x, x_g)$ is the minimal time to reach the goal from the start subject to passing through the state $x$.
%
%
%\subsection{Informed set}
Recall that for Euclidean spaces minimizing path length, the informed set~\Cinf is a prolate hyperspheroid~\cite{GSB14}.
Moreover, the size and shape of the hyperspheroid is defined only be the cost $c_{\text{best}}$ of the current best solution and not by the location of the start~$x_s$ and goal~$x_g$.

For the case of a MTDI, this is not the case. 
Specifically, we have that 
(i)~the structure of~\Cinf changes not only with~$c_{\text{best}}$ but also according to the specific values of~$x_s$ and~$x_g$ 
and that
(ii)~the cost map that implicitly defines~\Cinf can contain discontinuities (in contrast to Euclidean spaces minimizing path length where the cost map is continuous and differentiable at every point).

To understand the differences recall that optimal trajectories  for MTDI follow a ``bang-bang`` controller~\cite{HN10, KS14}.
Namely, we first apply maximal (or minimal) acceleration for some duration and then switch to applying minimal (or maximal, respectively) acceleration.
It is straightforward to see that both the type and the amount of acceleration applied (and hence the structure of~\Cinf) depend on the specific values of~$x_s$ and~$x_g$. 
%Fig~\ref{fig:informed_sets} depicts two informed sets defined using the same cost~$c_{\text{best}}$ but different start and goal states. Notice the difference in structure of~\Cinf. 
Fig~\ref{fig:discont} depicts a simple example where the cost map is discontinuous.

@@

In this section we describe our approach to efficiently produce new samples in an informed set~\Cinf given a specific cost~$c_{\text{best}}$ of trajectory $\gamma_{best}(t)$.
The samples follow a Markov Chain Monte Carlo, in which a new sample candidate is produced from a previous sample that also lies in the same informed set.
%However, this sample depends on the previous iterations of the algorithm which is why we describe it as a method to produce a sequence of samples.
Furthermore, the value~$c_{\text{best}}$ can decrease between consecutive iterations in the planning process of an informed RRT* planner.
This will occur if the search algorithm that uses the sampler finds a path to the goal whose cost is lower than~$c_{\text{best}}$.

The idea behind applying MCMC for informed sampling is to define a target distribution~$ \pi $ that has $p(x_{sample}\in\Cinf) \ne 0 ~\&~ p(x_{sample})\notin\Cinf = 0$. 
This is specially useful if we want to bias the samples based on our knowledge of the environment.
However, we make no such assumption about the environment and use a uniform distribution over all points in~$ \Cinf $. 
Our approach consists of two stages,
\begin{enumerate}
	\item finding an initial sample $ x_0 \in \Cinf $ which will serve as the start of a Markov chain. 
	This is implemented using the function 
	$ \texttt{sample\_in\_informed\_space}( )$, and
	\item sampling a new sample $ x_i \in \Cinf $ given a previous sample $ x_{i-1} $. 
	This is implemented using the function $\texttt{MCMC\_sample} (x_{i-1}, c_{\text{best}})$.
\end{enumerate}
Our framework is described in Alg.~\ref{alg:mcmc_informed_sampling} and visualized in Fig.~\ref{fig:alg}.
We now continue to detail each of the algorithm's stages.
%\os{Fix Alg 1 }

@@

Our approach is general and can be applied to any MCMC algorithm (see Sec.~\ref{sec:related_work}).
The process is demonstrated in Algorithm~\ref{alg:mcmc_informed_sampling}.
At the beginning of a Markov chain, \texttt{sample\_in\_informed\_space}() is called to generate the first sample in an informed set.
\texttt{MCMC\_sample}() is called to generate a new sample based on a previous sample $ x_{i-1} $ and a cost $ c_{best} $ that defines an informed set.
We demonstrate how to instantiate it with two different algorithms \emph{Metropolis-Hastings} and \emph{Hit-and-Run}, which will be described in later subsections.
If a generated new sample candidate is in the informed set, this candidate will be returned as a new sample.
But if a generated new sample candidate is not in the informed set, it will go back to line 2.
A new Markov chain will be initiated by calling \texttt{sample\_in\_informed\_space}() to generate a new sample $ x_0 $.

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\LOOP 
		\IF{ $ x_{i-1} == \varnothing $ }  
		\label{alg:mcmc_informed_sampling:start}
			\STATE $x_{0} \leftarrow \texttt{sample\_in\_informed\_space}( )$
		\ENDIF
		\STATE $x_{i} \leftarrow \texttt{MCMC\_sample} (x_{i-1}, c_{\text{best}})$ 
		\IF{ $ x_{i} \not\in \Cinf $ }
			\STATE $ i \leftarrow 0 $; $ x_{0} \leftarrow \varnothing $
			\STATE \textbf{Goto} line \ref{alg:mcmc_informed_sampling:start} 
		\ENDIF
		\STATE \textbf{return} $x_{i}$
		\ENDLOOP 
	\end{algorithmic}
	\caption{\captionstyle MCMC-based Informed Sampling $( x_{i-1}, c_{\text{best}})$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

@@

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $ x'_{i} \leftarrow \texttt{sample\_normal}( q ( x \mid x_{i-1},\Sigma) ) $ 
		\label{start}
		\STATE $ \alpha \leftarrow \frac{ q ( x_{i-1} \mid x'_{i},\Sigma) \pi( x'_{i} ) }{ q ( x'_{i} \mid x_{i-1},\Sigma) \pi( x_{i-1} ) } $
		\IF{ \texttt{sample\_random}$ (0.0 , 1.0) > \min (1, \alpha) $  }
            \RETURN $ x'_{i} $
		\ENDIF
		\RETURN $ x_{i} $
	\end{algorithmic}
	\caption{\captionstyle Metropolis-Hastings MCMC $(x_{i-1}, c_{\text{best}})$}
	\label{alg:mh_mcmc}	
\end{algorithm}

@@

		\IF{ \textsc{Cost}($ x_{i} $) $ < c_{\text{best}} $ }
		\RETURN $ x_{i}$
		\ENDIF

@@

		\ENDLOOP
	\end{algorithmic}
	\caption{\captionstyle Hit-And-Run MCMC $(x_{i-1}, c_{\text{best}})$}
	\label{alg:hit_and_run_mcmc}	
\end{algorithm}

@@

Figure \ref{fig:sampling_efficiency:levelset} shows how the informed set volume ratio decreases as the informed set cost $ c_{best} $ becomes smaller in problems of different dimensions.
In higher dimensions, the informed set volume ratio decreases much more quickly with decrease in the informed set cost $ c_{best} $, as new cheaper trajectories are found in the planning process.

@@

\begin{figure}[t!]
    \centering
	\includegraphics[width=.3\textwidth]{fig/sampling_efficiency/levelset}
	\caption{\captionstyle The informed set volume ratio decrease as $ c_{best} $ decreases in the planning process in different dimensions.}
	\label{fig:sampling_efficiency:levelset}
\end{figure}

\begin{figure}[t!]
	\centering
	\begin{subfigure}[b]{\linewidth}
		\centering
		\includegraphics[width=.70\linewidth]{fig/sampling_efficiency/sample_efficiency_2d}
		\caption{\captionstyle 4-dimension sampling space.}
		\label{fig:sampling_efficiency:2d}
	\end{subfigure}
	\begin{subfigure}[b]{\linewidth}
		\centering
		\includegraphics[width=.70\linewidth]{fig/sampling_efficiency/sample_efficiency_6d}
		\caption{\captionstyle 12-dimension sampling space.}
		\label{fig:sampling_efficiency:6d}
	\end{subfigure}
	\caption{\captionstyle Average sampling time vs informed set volume ratio of four samplers (RS, HRS, MH and HNR) in state spaces of different dimensions. 
	The X axis is the ratio of informed set volume to the entire sampling space.
	The Y axis is the average time per sample.}
	\label{fig:sampling_efficiency}
\end{figure}

@@

The quality of samples determines the efficiency of resulting planning algorithms.
If a sampler could provide samples with same quality as others but generate samples in a much efficiency way, we would expect that an informed RRT* with this sampler would show two properties.
\begin{itemize}
	\item It shall converge faster in finding the optimal solution.
	Sampling in an informed set is gradually becoming harder as new better solution reduces $ c_{best} $ which reduces the informed set volume ratio.
	\item Its performance should not degrade significantly in high dimensional problems.
	As shown in Figure \ref{fig:sampling_efficiency:levelset}, the informed set volume ratio decrease more significantly in a high-dimension state space.
	The advantage of a good informed sampler becomes evident.
\end{itemize}

@@

\begin{figure*}[t!]
	\centering
	\begin{subfigure}[b]{0.325\textwidth}
		\includegraphics[width=\linewidth]{fig/planning_efficiency/3dof_general}
		\caption{\captionstyle Problem 1 - 6 Dimensions.}
		\label{fig:planning_efficiency:3dof:general}
	\end{subfigure}	
	\begin{subfigure}[b]{0.325\textwidth}
		\includegraphics[width=\linewidth]{fig/planning_efficiency/6dof_hammering}
		\caption{\captionstyle Problem 2 - 12 Dimensions.}
		\label{fig:planning_efficiency:6dof:hammering}
	\end{subfigure}
	\begin{subfigure}[b]{0.325\textwidth}
	\includegraphics[width=\linewidth]{fig/planning_efficiency/herb_batting_efficiency}
	\caption{\captionstyle Problem 3 - 14 Dimensions.}
	\label{fig:planning_efficiency:herb:batting}
    \end{subfigure}
	\caption{\captionstyle Planning Efficiency of four different samplers (RS, HRS, MH and HNR) in three problems. 
	The X axis is the planning time. 
	The Y axis is the ration of the current best and the optimal $ c_{best} / c^*_{best} $.}
	\label{fig:planning_efficiency}
\end{figure*}

>>>>

Given a trajectory $\gamma_{\rm best}$ with cost $\cbest = c(\gamma_{\rm best})$ the \emph{informed set}~\Cinf is defined to be all states $x$  which may be on trajectories with lower cost than $\cbest$.
Specifically,
$
\Cinf = \{ x \in \calX \mid  
		c ( \gamma^*(x) ) < \cbest \} $~\cite{GSB14}.
Here~$ \gamma^*(x) $ denotes the optimal trajectory  from $ x_s $ to $ x_g $ constrained to pass through $ x $.
Notice that we do not require that~$ \gamma^*(x) $ is collision free.

@@

%Let $T(x_1, x_2)$ denote the minimal time to move the robot from state $x_1$ to $x_2$.
%Furthermore, set
%%
%\begin{equation}
%T(x_s, x, x_g) = T(x_s, x) + T(x, x_g).
%\end{equation}
%%
%Namely, $T(x_s, x, x_g)$ is the minimal time to reach the goal from the start subject to passing through the state $x$.
%
%
%\subsection{Informed set}
Recall that for Euclidean spaces minimizing path length, the informed set~\Cinf is a prolate hyperspheroid~\cite{GSB14}.
Moreover, the size and shape of the hyperspheroid is defined only be the cost $\cbest$ of the current best solution and not by the location of the start~$x_s$ and goal~$x_g$.

For the case of a MTDI, this is not the case. 
Specifically, we have that 
(i)~the structure of~\Cinf changes not only with~$\cbest$ but also according to the specific values of~$x_s$ and~$x_g$ 
and that
(ii)~the cost map that implicitly defines~\Cinf can contain discontinuities (in contrast to Euclidean spaces minimizing path length where the cost map is continuous and differentiable at every point).

To understand the differences recall that optimal trajectories  for MTDI follow a ``bang-bang`` controller~\cite{HN10, KS14}.
Namely, we first apply maximal (or minimal) acceleration for some duration and then switch to applying minimal (or maximal, respectively) acceleration.
It is straightforward to see that both the type and the amount of acceleration applied (and hence the structure of~\Cinf) depend on the specific values of~$x_s$ and~$x_g$. 
%Fig~\ref{fig:informed_sets} depicts two informed sets defined using the same cost~$\cbest$ but different start and goal states. Notice the difference in structure of~\Cinf. 
Fig~\ref{fig:discont} depicts a simple example where the cost map is discontinuous.

@@

In this section we describe our approach to efficiently produce new samples in an informed set~\Cinf given a specific cost~$\cbest$ of trajectory $\gamma_{\rm best}(t)$.
The samples follow a Markov Chain Monte Carlo, in which a new sample candidate is produced from a previous sample that also lies in the same informed set.
%However, this sample depends on the previous iterations of the algorithm which is why we describe it as a method to produce a sequence of samples.
Furthermore, the value~$\cbest$ can decrease between consecutive iterations in the planning process of an informed RRT* planner.
This will occur if the search algorithm that uses the sampler finds a path to the goal whose cost is lower than~$\cbest$.

The idea behind applying MCMC for informed sampling is to define a target distribution~$ \pi $ that has $p(x_{sample}\in\Cinf) \ne 0 ~\&~ p(x_{sample})\notin\Cinf = 0$. 
This is specially useful if we want to bias the samples based on our knowledge of the environment.
However, we make no such assumption about the environment and use a uniform distribution over all points in~$ \Cinf $. 
Our approach consists of two stages,
\begin{enumerate}
	\item finding an initial sample $ x_0 \in \Cinf $ which will serve as the start of a Markov chain. 
	This is implemented using the function 
	$ \texttt{sample\_in\_informed\_space}( )$, and
	\item sampling a new sample $ x_i \in \Cinf $ given a previous sample $ x_{i-1} $. 
	This is implemented using the function $\texttt{MCMC\_sample} (x_{i-1}, \cbest)$.
\end{enumerate}
Our framework is described in Alg.~\ref{alg:mcmc_informed_sampling} and visualized in Fig.~\ref{fig:alg}.
We now continue to detail each of the algorithm's stages.
%\os{Fix Alg 1 }

@@

Our approach is general and can be applied to any MCMC algorithm (see Sec.~\ref{sec:related_work}).
The process is demonstrated in Algorithm~\ref{alg:mcmc_informed_sampling}.
At the beginning of a Markov chain, \texttt{sample\_in\_informed\_space}() is called to generate the first sample in an informed set.
\texttt{MCMC\_sample}() is called to generate a new sample based on a previous sample $ x_{i-1} $ and a cost $ \cbest $ that defines an informed set.
We demonstrate how to instantiate it with two different algorithms \emph{Metropolis-Hastings} and \emph{Hit-and-Run}, which will be described in later subsections.
If a generated new sample candidate is in the informed set, this candidate will be returned as a new sample.
But if a generated new sample candidate is not in the informed set, it will go back to line 2.
A new Markov chain will be initiated by calling \texttt{sample\_in\_informed\_space}() to generate a new sample $ x_0 $.

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\LOOP 
		\IF{ $ x_{i-1} == \varnothing $ }  
		\label{alg:mcmc_informed_sampling:start}
			\STATE $x_{0} \leftarrow \texttt{sample\_in\_informed\_space}( )$
		\ENDIF
		\STATE $x_{i} \leftarrow \texttt{MCMC\_sample} (x_{i-1}, \cbest)$ 
		\IF{ $ x_{i} \not\in \Cinf $ }
			\STATE $ i \leftarrow 0 $; $ x_{0} \leftarrow \varnothing $
			\STATE \textbf{Goto} line \ref{alg:mcmc_informed_sampling:start} 
		\ENDIF
		\STATE \textbf{return} $x_{i}$
		\ENDLOOP 
	\end{algorithmic}
	\caption{\captionstyle MCMC-based Informed Sampling $( x_{i-1}, \cbest)$}
	\label{alg:mcmc_informed_sampling}	
\end{algorithm}

@@

\begin{algorithm}[t]
	\begin{algorithmic}[1]
		\STATE $ x'_{i} \leftarrow \texttt{sample\_normal}( q ( x \mid x_{i-1},\Sigma) ) $ 
		\label{start}
		\STATE $ \alpha \leftarrow \frac{ q ( x_{i-1} \mid x'_{i},\Sigma) \pi( x'_{i} ) }{ q ( x'_{i} \mid x_{i-1},\Sigma) \pi( x_{i-1} ) } $
		\IF{ \texttt{sample\_random}$ (0.0 , 1.0) > \min (1, \alpha) $  }
            \RETURN $ x'_{i} $
		\ENDIF
		\RETURN $ x_{i} $
	\end{algorithmic}
	\caption{\captionstyle Metropolis-Hastings MCMC $(x_{i-1}, \cbest)$}
	\label{alg:mh_mcmc}	
\end{algorithm}

@@

		\IF{ \textsc{Cost}($ x_{i} $) $ < \cbest $ }
		\RETURN $ x_{i}$
		\ENDIF

@@

		\ENDLOOP
	\end{algorithmic}
	\caption{\captionstyle Hit-And-Run MCMC $(x_{i-1}, \cbest)$}
	\label{alg:hit_and_run_mcmc}	
\end{algorithm}

@@

Figure \ref{fig:sampling_efficiency:levelset} shows how the informed set volume ratio decreases as the informed set cost $ \cbest $ becomes smaller in problems of different dimensions.
In higher dimensions, the informed set volume ratio decreases much more quickly with decrease in the informed set cost $ \cbest $, as new cheaper trajectories are found in the planning process.

@@

\begin{figure*}[t!]
	\centering
	\begin{subfigure}[t!]{0.325\textwidth}
		\centering
		\includegraphics[width=\textwidth]{fig/sampling_efficiency/levelset}
		\caption{\captionstyle The informed set volume ratio decrease as $ \cbest $ decreases in the planning process in different dimensions.}
		\label{fig:sampling_efficiency:levelset}
	\end{subfigure}
	\begin{subfigure}[t!]{0.325\textwidth}
		\centering
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_2d}
		\caption{\captionstyle 4-dimension sampling space.}
		\label{fig:sampling_efficiency:2d}
	\end{subfigure}
	\begin{subfigure}[t!]{0.325\textwidth}
		\centering
		\includegraphics[width=\linewidth]{fig/sampling_efficiency/sample_efficiency_6d}
		\caption{\captionstyle 12-dimension sampling space.}
		\label{fig:sampling_efficiency:6d}
	\end{subfigure}
	\caption{\captionstyle Average sampling time vs informed set volume ratio of four samplers (RS, HRS, MH and HNR) in state spaces of different dimensions. 
	The X axis is the ratio of informed set volume to the entire sampling space.
	The Y axis is the average time per sample.}
	\label{fig:sampling_efficiency}
\end{figure*}

@@

The quality of samples determines the efficiency of resulting planning algorithms.
If a sampler could provide samples with same quality as others but generate samples in a much efficiency way, we would expect that an informed RRT* with this sampler would show two properties.
\begin{itemize}
	\item It shall converge faster in finding the optimal solution.
	Sampling in an informed set is gradually becoming harder as new better solution reduces $ \cbest $ which reduces the informed set volume ratio.
	\item Its performance should not degrade significantly in high dimensional problems.
	As shown in Figure \ref{fig:sampling_efficiency:levelset}, the informed set volume ratio decrease more significantly in a high-dimension state space.
	The advantage of a good informed sampler becomes evident.
\end{itemize}

@@

\begin{figure*}[t!]
	\centering
	\begin{subfigure}[b]{0.325\textwidth}
		\includegraphics[width=\linewidth]{fig/planning_efficiency/3dof_general}
		\caption{\captionstyle Problem 1 - 6 Dimensions.}
		\label{fig:planning_efficiency:3dof:general}
	\end{subfigure}	
	\begin{subfigure}[b]{0.325\textwidth}
		\includegraphics[width=\linewidth]{fig/planning_efficiency/6dof_hammering}
		\caption{\captionstyle Problem 2 - 12 Dimensions.}
		\label{fig:planning_efficiency:6dof:hammering}
	\end{subfigure}
	\begin{subfigure}[b]{0.325\textwidth}
	\includegraphics[width=\linewidth]{fig/planning_efficiency/herb_batting_efficiency}
	\caption{\captionstyle Problem 3 - 14 Dimensions.}
	\label{fig:planning_efficiency:herb:batting}
    \end{subfigure}
	\caption{\captionstyle Planning Efficiency of four different samplers (RS, HRS, MH and HNR) in three problems. 
	The X axis is the planning time. 
	The Y axis is the ration of the current best and the optimal $ \cbest / c^*_{\rm best} $.}
	\label{fig:planning_efficiency}
\end{figure*}
